---
title: "Coursera PML Project"
author: "Ramandha Putra S"
date: "3/4/2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(randomForest)
library(gbm)
library(dplyr)
library(ggplot2)
```

# Introduction

This project aims to predict the manner in which individuals perform barbell lifts using accelerometer data. The target variable is `classe`, which represents different exercise types.

# Data Preprocessing

```{r load-data}
# Load Data
train_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
train_data <- read.csv(train_url, na.strings = c("NA", "", "#DIV/0!"))
test_data <- read.csv(test_url, na.strings = c("NA", "", "#DIV/0!"))

# Remove Metadata Columns (First 5 columns)
train_data <- train_data[, -(1:5)]
test_data <- test_data[, -(1:5)]

# Remove Columns with Too Many NAs
threshold <- 0.5 * nrow(train_data)
valid_cols <- colSums(is.na(train_data)) < threshold
train_data <- train_data[, valid_cols]
test_data <- test_data[, valid_cols]
```

# Train-Test Split

```{r split-data}
set.seed(123)
inTrain <- createDataPartition(train_data$classe, p = 0.75, list = FALSE)
training <- train_data[inTrain, ]
testing <- train_data[-inTrain, ]

# Ensure Target Variable is a Factor
training$classe <- as.factor(training$classe)
testing$classe <- as.factor(testing$classe)
```

# Model Training

```{r train-models}
set.seed(123)
ctrl <- trainControl(method = "cv", number = 5)

# Random Forest Model
rf_model <- train(classe ~ ., data = training, method = "rf", trControl = ctrl)

# Gradient Boosting Model
gbm_model <- train(classe ~ ., data = training, method = "gbm", trControl = ctrl, verbose = FALSE)
```

# Model Evaluation

```{r evaluate-models}
pred_rf <- predict(rf_model, testing)
pred_gbm <- predict(gbm_model, testing)

conf_rf <- confusionMatrix(pred_rf, testing$classe)
conf_gbm <- confusionMatrix(pred_gbm, testing$classe)

print(conf_rf)
print(conf_gbm)
```

# Prediction on Test Data

```{r predict-test-data}
predictions <- predict(rf_model, test_data)
write.csv(predictions, "final_predictions.csv", row.names = FALSE)
```

# Conclusion

The Random Forest model achieved high accuracy, making it a suitable choice for predicting exercise classification. The model will be used to predict the test dataset for final submission.